---
title: "401 project"
author: "JACK LOHMAN"
date: "2025-03-12"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
library(lubridate)
library(survival)
library(survminer)
library(caret)   
library(nnet)
library(data.table)
dispatch_data <- read_csv("big_dispatch.csv")
```


```{r}
# Data cleaning

dispatch_data <- dispatch_data %>%
  select(-source_filename, -data_loaded_at, -data_updated_at, -dup_cad_number, 
         -analysis_neighborhood, -police_district, -supervisor_district, -intersection_name, -intersection_id,
         -intersection_point, ) %>%
  filter(
      !(call_type_final_desc %in% c("EMERGENCY BACKUP", "CITIZEN ARREST", "RESISTING ARREST", 
                      "URGENT NOTIFICATION", "YGC / JAIL BREAK","FIRE", "EXPLOSION", "EXPLOSIVE FOUND", "HAZMAT INCIDENT", 
                      "DRUNK DRIVER", "ROLLING DRUNK PERSON", "INJURY VEH ACCIDENT", 
                      "BLOCKED SIDEWALK", "ALARM", "PANIC ALARM", "BOMB THREAT", "SUSPICIOUS MAILING")),  
   
      !(disposition %in% c(NA, "NA", "UNKNOWN", "CANCELLED", "GOA", "UTL")),
      !(onscene_datetime %in% c(NA, "NA")),
      !(close_datetime %in% c(NA, "NA"))
      )
  
dispatch_data <- dispatch_data %>%
  mutate(
    onscene_datetime = as.POSIXct(onscene_datetime, format="%Y/%m/%d %I:%M:%S %p"),
    close_datetime = as.POSIXct(close_datetime, format="%Y/%m/%d %I:%M:%S %p"),
    enroute_datetime = as.POSIXct(enroute_datetime, format="%Y/%m/%d %I:%M:%S %p"),
    mins_onscene_to_close = as.numeric(difftime(close_datetime, onscene_datetime, units = "mins"))
  )
    
unique(dispatch_data$call_type_final_desc)



```

                      
```{r}
dispatch_data <- dispatch_data %>%
  filter(enroute_datetime > as.POSIXct("2022-01-01 00:00:00", format="%Y-%m-%d %H:%M:%S"))

dispatch_data <- dispatch_data %>%
  mutate(incedent_cat = case_when(

    call_type_final_desc %in% c("FIGHT NO WEAPON DV", "ASSAULT / BATTERY", "ASSAULT / BATTERY DV", "ASSAULT / BATTERY EA", 
                      "ASSAULT / BATTERY CA", "AGG ASSAULT / ADW", "AGG ASSAULT / ADW DV", 
                      "AGG ASSAULT / ADW CA", "AGG ASSAULT / ADW EA", "SEXUAL ASSAULT ADULT", 
                      "SEXUAL ASSAULT JUVE", "HOMICIDE", "STABBING", "STABBING DV", 
                      "SHOOTING", "SHOOTING DV", "SHOTS FIRED", "KIDNAPPING", "PERSON BREAK IN DV", 
                      "FIGHT NO WEAPON") ~ "Violent Crimes",
    
    # Property Crimes
    call_type_final_desc %in% c("PETTY THEFT", "PETTY THEFT EA", "GRAND THEFT", "GRAND THEFT EA", 
                      "AUTO BOOST / STRIP", "STOLEN PROPERTY", "FRAUD", "FRAUD EA", 
                      "BURGLARY", "ROBBERY", "STRONGARM ROBBERY", "PURSE SNATCH", 
                      "STOLEN VEHICLE", "WANTED VEHICLE / SUB", "PERSON BREAKING IN") ~ "Property Crimes",
    
    # Disturbances & Public Order
    call_type_final_desc %in% c("NOISE NUISANCE", "DEMO / PROTEST", "TRESPASSER", "SUSPICIOUS PERSON", 
                      "SUSPICIOUS VEHICLE", "SURVEILLANCE", "PERSON W/KNIFE", "PERSON W/KNIFE DV", 
                      "PERSON W/KNIFE CRIT", "PERSON W/GUN", "PERSON W/GUN DV", 
                      "THREATS / HARASSMENT", "THREATS DV", "THREATS EA", 
                      "STALKING", "STALKING DV") ~ "Disturbances & Public Order",
    
    # Mental Health & Welfare Checks
    call_type_final_desc %in% c("MENTALLY DISTURBED", "MENTALLY DIST CRIT", "PSYCH EVAL / HOLD", 
                      "SUICIDE ATTEMPT", "SUICIDE ATT CRIT", "WELL BEING CHECK", 
                      "WELL BEING CHECK DV", "ELDERLY ABUSE", "SENILE PERSON", 
                      "MISSING ADULT", "MISSING JUVENILE", "JUVE BEYOND CONTROL", 
                      "JUVENILE DISTURBANCE") ~ "Mental Health & Welfare Checks",
    
    

    TRUE ~ "Other/Uncategorized"
  ))
```


```{r}
setDT(dispatch_data)

dispatch_data[, call_volume := 
  dispatch_data[.SD, 
    on = .(enroute_datetime < close_datetime, close_datetime > enroute_datetime), 
    .N, by = .EACHI]$N - 1
]

dispatch_data <- as.data.frame(dispatch_data)

dispatch_data


```


```{r}
dispatch_data <- dispatch_data %>%
  mutate(punitive_action = case_when(

    disposition %in% c("ABA", "ADM", "ADV", "CAN", "CSA", "HAN", "PAS", "NOM", "ND", "NCR", "VAS") ~ 0,
    
    disposition %in% c("ARR", "CIT", "CRM", "REP") ~ 1,
    
    TRUE ~ NA
  ))


dispatch_data

```

```{r}
set.seed(123)  
trainIndex <- createDataPartition(iris$Species, p = 0.8, list = FALSE)
trainData <- dispatch_data[trainIndex, ]
testData <- dispatch_data[-trainIndex, ]

```


```{r}
set.seed(123)

train_control <- trainControl(method = "cv", number = 5)  # 5-fold cross-validation



multi_logit_model <- train(incedent_cat ~ mins_onscene_to_close + call_volume +
                           data = trainData, 
                           method = "multinom", 
                           trControl = train_control)

# Print model summary
print(multi_logit_model)

```


```{r}
# Predict on test data
predictions <- predict(multi_logit_model, newdata = testData)

# Confusion Matrix to evaluate performance
conf_matrix <- confusionMatrix(predictions, testData$Species)
print(conf_matrix)

```

ABA
Abated
Officer responded, confirmed the call, and handled
ADM
Admonished
Officer responded, confirmed the call, and handled
ADV
Advised
Officer responded, confirmed the call, and handled
ARR
Arrest
Officer indicates an arrest. Note: Will be different from other SFPD reporting and is should not be used to calculate arrest statistics
CAN
Cancel
Call cancelled before officers engaged the call
CSA
CPSA assignment

CPSA assignment
22
Cancel
Call cancelled before officers engaged the call
CIT
Cited
Citation was issued
CRM
Criminal Activation
Burglary Alarm calls resulting in criminal activity
GOA
Gone on Arrival
Upon arriving suspect no longer on scene.
HAN
Handled
Officer responded, confirmed the call, and handled
NCR
Non-Criminal
Arrived on scene and did not find a criminal issue
ND
No Disposition
Officer determined this call related to another call. The other call would have the disposition for the incident.
NOM
No Merit
Arrive on scene and Officer determines doesn’t merit further response
PAS
Premises Appears Secure
Premises Appears Secure. Disposition commonly seen for house alarm calls.
REP
Report
Officer resolved incident and this resulted in a police report. Note: Should not be used to calculate official number of reports
SFD
SFFD Medical Staff Engaged
EMS or SFFD Medical Staff Engaged
UTL
Unable to Locate
Upon arriving suspect no longer on scene
VAS